# Network analysis

```{r load_overall}
order_colors <- c(
  "#08D1D1","#BF7022","#2EBC82","#2E91BC","#2E5ABC","#09CE8D","#E0A608","#097265","#AB56D8","#0AA01F","#274945","#4E397A","#B3265B",
  "#2EBC45","#7F4561","#58BC2E","#BC2EA1","#6A52C6","#AD8D8D", "#CFE71F")

# order_colors <- c(
#   "#08D1D1","#BF7022","#2EBC82","#2E91BC","#2E5ABC","#09CE8D","#E0A608","#097265","#AB56D8","#0AA01F","#274945","#4E397A","#B3265B",
#   "#2EBC45","#7F4561","#58BC2E","#BC2EA1","#6A52C6","#AD8D8D")

```


## Overall 


```{r overvall}
cxFile  <- "microbetag_analysis/microbetag_nets/overall.cx"
overall = process_cxfile(cxFile, "overall", genome_metadata, genome_counts_filt, order_colors)
```

```{r plots-overall}
all_days_pos_plot <- plot_graph(overall$graph_pos, overall$clusters_pos, "all_days_pos.png")
all_days_neg_plot <- plot_graph(overall$graph_neg, overall$clusters_neg, "all_days_neg.png")
```


## Timewise analysis

### Day 7

```{r cor_day7, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_day_7.cx"
day7 = process_cxfile(cxFile, day, "day7", genome_metadata, genome_counts_filt, order_colors)
```


```{r cor_day7_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
d7_pos_plot <- plot_graph(day7$graph_pos, day7$clusters_pos, "day07_pos.png")
d7_neg_plot <- plot_graph(day7$graph_neg, day7$clusters_neg, "day07_neg.png")
```

### Day 14

```{r cor_day14, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_day_14.cx"
day14 = process_cxfile(cxFile, day, "day14", genome_metadata, genome_counts_filt, order_colors)
```

```{r cor_day14_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
d14_pos_plot <- plot_graph(day14$graph_pos, day14$clusters_pos, "day14_pos.png")
d14_neg_plot <- plot_graph(day14$graph_neg, day14$clusters_neg, "day14_neg.png")
```

### Day 21

```{r cor_day21, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_day_21.cx"
day21 = process_cxfile(cxFile, day, "day21", genome_metadata, genome_counts_filt, order_colors)
```


```{r cor_day21_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
d21_pos_plot <- plot_graph(day21$graph_pos, day21$clusters_pos, "day21_pos.png")
d21_neg_plot <- plot_graph(day21$graph_neg, day21$clusters_neg, "day21_neg.png")
```

### Day 28

```{r cor_day28, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_day_28.cx"
day28 = process_cxfile(cxFile, day, "day28", genome_metadata, genome_counts_filt, order_colors)
```


```{r cor_day28_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
d28_pos_plot <- plot_graph(day28$graph_pos, day28$clusters_pos, "day28_pos.png")
d28_neg_plot <- plot_graph(day28$graph_neg, day28$clusters_neg, "day28_neg.png")
```

### Day 35

```{r cor_day35, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_day_35.cx"
day35  <- process_cxfile(cxFile, day, "day35", genome_metadata, genome_counts_filt, order_colors)
```


```{r cor_day35_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
d35_pos_plot <- plot_graph(day35$graph_pos, day35$clusters_pos, "day35_pos.png")
d35_neg_plot <- plot_graph(day35$graph_neg, day35$clusters_neg, "day35_neg.png")
```



```{r as}
# POSITIVE
img1 <- readPNG("day07_pos.png")
img2 <- readPNG("day14_pos.png")
img3 <- readPNG("day21_pos.png")
img4 <- readPNG("day28_pos.png")
img5 <- readPNG("day35_pos.png")

# NEGATIVE
img6 <- readPNG("day07_neg.png")
img7 <- readPNG("day14_neg.png")
img8 <- readPNG("day21_neg.png")
img9 <- readPNG("day28_neg.png")
img10 <- readPNG("day35_neg.png")


img_list <- list(img1, img2, img3, img4, img5, img6, img7, img8, img9, img10)

png("grid_plot.png", width = 5000, height = 1000, res = 300)  # Adjust size and resolution as needed


grid_objects <- lapply(img_list, function(img) {
  rasterGrob(img, interpolate = TRUE)
})

grid.arrange(grobs = grid_objects, ncol = 5, nrow = 2)
grid.text("positive", x = unit(0.03, "npc"), y = unit(0.75, "npc"), rot = 90, gp = gpar(fontsize = 14))
grid.text("negative", x = unit(0.03, "npc"), y = unit(0.25, "npc"), rot = 90, gp = gpar(fontsize = 14))

dev.off()

```




## Treatment-wise analysis

### TG1

```{r cor_TG1, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_TG1.cx"
TG1  <- process_cxfile(cxFile, treatment, "TG1", genome_metadata, genome_counts_filt, order_colors)
```


```{r cor_TG1_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
TG1_pos_plot <- plot_graph(TG1$graph_pos, TG1$clusters_pos, "TG1_pos.png")
TG1_neg_plot <- plot_graph(TG1$graph_neg, TG1$clusters_neg, "TG1_neg.png")
```

### TG2

```{r cor_TG2, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_TG2.cx"
TG2    <- process_cxfile(cxFile, treatment, "TG2", genome_metadata, genome_counts_filt, order_colors)
```

```{r cor_TG2_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
TG2_pos_plot <- plot_graph(TG2$graph_pos, TG2$clusters_pos, "TG2_pos.png")
TG2_neg_plot <- plot_graph(TG2$graph_neg, TG2$clusters_neg, "TG2_neg.png")
```

### TG3

```{r cor_TG3, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_TG3.cx"
TG3    <- process_cxfile(cxFile, treatment, "TG3", genome_metadata, genome_counts_filt, order_colors)
```

```{r cor_TG3_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
TG3_pos_plot <- plot_graph(TG3$graph_pos, TG3$clusters_pos, "TG3_pos.png")
TG3_neg_plot <- plot_graph(TG3$graph_neg, TG3$clusters_neg, "TG3_neg.png")
```

### TG4

```{r cor_TG4, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_TG4.cx"
TG4    <- process_cxfile(cxFile, treatment, "TG4", genome_metadata, genome_counts_filt, order_colors)
```


```{r cor_TG4_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
TG4_pos_plot <- plot_graph(TG4$graph_pos, TG4$clusters_pos, "TG4_pos.png")
TG4_neg_plot <- plot_graph(TG4$graph_neg, TG4$clusters_neg, "TG4_neg.png")
```

### TG5

```{r cor_TG5, comment="", message=FALSE, warning=FALSE}
cxFile <- "microbetag_analysis/microbetag_nets/microbetag_net_TG5.cx"
TG5    <- process_cxfile(cxFile, treatment, "TG5", genome_metadata, genome_counts_filt, order_colors)
```

```{r cor_TG5_clusters, message=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
TG5_pos_plot <- plot_graph(TG5$graph_pos, TG5$clusters_pos, "TG5_pos.png")
TG5_neg_plot <- plot_graph(TG5$graph_neg, TG5$clusters_neg, "TG5_neg.png")
```

## Combined

### Network properties

How the nodes in the graph tend to cluster together, indicating how tightly correlated groups of genomes are.

```{r network_properties, comment="", message=FALSE, warning=FALSE}
bind_rows(day7$metrics_pos, 
          day14$metrics_pos, 
          day21$metrics_pos,
          day28$metrics_pos,
          day35$metrics_pos) %>% 
  tt()

bind_rows(day7$metrics_neg, 
          day14$metrics_neg, 
          day21$metrics_neg,
          day28$metrics_neg,
          day35$metrics_neg) %>% 
  tt()

bind_rows(TG1$metrics_pos,
          TG2$metrics_pos,
          TG3$metrics_pos,
          TG4$metrics_pos,
          TG5$metrics_pos)%>%
  tt()

bind_rows(TG1$metrics_neg,
          TG2$metrics_neg,
          TG3$metrics_neg,
          TG4$metrics_neg,
          TG5$metrics_neg,) %>%
  tt()
```

```{r transitivity_plots, comment="", message=FALSE, warning=FALSE}
bind_rows(day7$metrics_pos, 
          day14$metrics_pos, 
          day21$metrics_pos, 
          day28$metrics_pos, 
          day35$metrics_pos) %>% 
  mutate(day=factor(day,levels=c("day7","day14","day21","day28","day35"))) %>%
  pivot_longer(!day, names_to="metric", values_to = "value") %>% 
  filter(metric != "assortability") %>% 
  ggplot(aes(x=day, y=value)) + 
    geom_col() +
    facet_wrap(. ~ metric, scales="free", nrow = 1) +
    theme_classic()

bind_rows(day7$metrics_neg, 
          day14$metrics_neg, 
          day21$metrics_neg, 
          day28$metrics_neg, 
          day35$metrics_neg) %>% 
  mutate(day=factor(day, levels=c("day7","day14","day21","day28","day35"))) %>%
  pivot_longer(!day, names_to="metric", values_to = "value") %>% 
  filter(metric != "assortability") %>%  
  ggplot(aes(x=day, y=value)) + 
    geom_col() +
    facet_wrap(. ~ metric, scales="free", nrow = 1) +
    theme_classic()

bind_rows(TG1$metrics_pos,
          TG2$metrics_pos,
          TG3$metrics_pos,
          TG4$metrics_pos,
          TG5$metrics_pos) %>%
  mutate(treatment=factor(treatment,levels=c("TG1","TG2","TG3","TG4","TG5"))) %>%
  pivot_longer(!treatment, names_to="metric", values_to = "value") %>%
  filter(metric != "assortability") %>%
  ggplot(aes(x=treatment, y=value)) +
    geom_col() +
    facet_wrap(. ~ metric, scales="free", nrow = 1) +
    theme_classic()

bind_rows(TG1$metrics_neg,
          TG2$metrics_neg,
          TG3$metrics_neg,
          TG4$metrics_neg,
          TG5$metrics_neg) %>%
  mutate(treatment=factor(treatment,levels=c("TG1","TG2","TG3","TG4","TG5"))) %>%
  pivot_longer(!treatment, names_to="metric", values_to = "value") %>%
  filter(metric != "assortability") %>%
  ggplot(aes(x=treatment, y=value)) +
    geom_col() +
    facet_wrap(. ~ metric, scales="free", nrow = 1) +
    theme_classic()
```

### Positive correlations

#### Clusters per day

```{r community_clusters_day, comment="", message=FALSE, warning=FALSE}
day_clusters_positive <- c(day7$communities_pos, 
                           day14$communities_pos,
                           day21$communities_pos,
                           day28$communities_pos,
                           day35$communities_pos) %>%
  purrr::map(~ combn(.x, 2, simplify = FALSE)) %>%
  unlist(recursive = FALSE) %>% 
  purrr::map(~ sort(.x))

# Calculate functional distances
day_clusters_positive_dist <- compute_cluster_distances(day_clusters_positive, genome_gifts)

# Print top pairs
day_clusters_positive_dist %>% 
  filter(count>2) %>% 
  tt()
```

#### Clusters per treatment

```{r community_clusters_treatment, comment="", message=FALSE, warning=FALSE}
treatment_clusters_positive <- c(TG1$communities_pos, 
                        TG2$communities_pos, 
                        TG3$communities_pos, 
                        TG4$communities_pos, 
                        TG5$communities_pos) %>%
  purrr::map(~ combn(.x, 2, simplify = FALSE)) %>%
  unlist(recursive = FALSE) %>% 
  purrr::map(~ sort(.x))

# Calculate functional distances
treatment_clusters_positive_dist <- compute_cluster_distances(treatment_clusters_positive, genome_gifts)

# Print top pairs
treatment_clusters_positive %>% 
  filter(count>3) %>% 
  tt()
```

#### Cluster combined

```{r community_clusters_all, comment="", message=FALSE, warning=FALSE}
all_clusters_positive <- c(day7$communities_pos, 
                           day14$communities_pos,
                           day21$communities_pos,
                           day28$communities_pos,
                           day35$communities_pos
                        ) %>%
  purrr::map(~ combn(.x, 2, simplify = FALSE)) %>%
  unlist(recursive = FALSE) %>% 
  purrr::map(~ sort(.x))

# Calculate functional distances
all_clusters_positive_dist <- compute_cluster_distances(all_clusters_positive, genome_gifts)

# Print top pairs
all_clusters_positive_dist %>% 
  filter(count>2) %>% 
  tt()
```

```{r community_clusters_all_plot, comment="", message=FALSE, warning=FALSE}
day_clusters_positive_filtered <- day_clusters_positive_dist %>% 
  filter(count>3) %>%
  rename(count_day=count) %>% 
  mutate(clustering_day=TRUE)

treatment_clusters_positive_filtered <- treatment_clusters_positive_dist %>% 
  filter(count>3) %>%
  rename(count_treatment=count) %>% 
  mutate(clustering_treatment=TRUE)

all_clusters_positive_filtered <- all_clusters_positive_dist %>% 
  filter(count > 5) %>%
  rename(count_all=count) %>% 
  mutate(clustering_all=TRUE)

associated_pairs_positive <- all_clusters_positive_filtered %>% 
  full_join(day_clusters_positive_filtered, by=join_by("genomes"=="genomes","distance"=="distance")) %>% 
  full_join(treatment_clusters_positive_filtered, by=join_by("genomes"=="genomes","distance"=="distance")) %>%
  rowwise() %>% 
  mutate(count=max(count_all,count_day,count_treatment, na.rm=TRUE)) %>%
  mutate(clustering = case_when(
    clustering_all == TRUE ~ "all",  # If clustering_all is TRUE
    clustering_day == TRUE & is.na(clustering_all) & is.na(clustering_treatment) ~ "day", 
    clustering_treatment == TRUE & is.na(clustering_all) & is.na(clustering_day) ~ "treatment")) %>%
  select(genomes,count,clustering,distance) %>%
  mutate(clustering=factor(clustering)) %>%
  #add coordinates
  separate(genomes, into = c("genome1", "genome2"), sep = " - ") %>%
  left_join(gift_pcoa_vectors %>% rownames_to_column(var = "genome"), by = c("genome1" = "genome")) %>%
  rename(x1 = Axis.1, y1 = Axis.2) %>%
  left_join(gift_pcoa_vectors %>% rownames_to_column(var = "genome"), by = c("genome2" = "genome")) %>%
  rename(x2 = Axis.1, y2 = Axis.2)

associated_pairs_positive %>% 
  select(genome1,genome2,count,clustering) %>% 
  left_join(genome_metadata %>% select(genome,order,species),by=join_by(genome1==genome)) %>% 
  left_join(genome_metadata %>% select(genome,order,species),by=join_by(genome2==genome)) %>% 
  tt()

```

### Negative correlations

#### Clusters per day

```{r community_clusters_day_negative, comment="", message=FALSE, warning=FALSE}
day_clusters_negative <- c(day7$communities_neg, 
                           day14$communities_neg,
                           day21$communities_neg,
                           day28$communities_neg,
                           day35$communities_neg) %>%
  purrr::map(~ combn(.x, 2, simplify = FALSE)) %>%
  unlist(recursive = FALSE) %>% 
  purrr::map(~ sort(.x))

# Calculate functional distances
day_clusters_negative_dist <- compute_cluster_distances(day_clusters_negative, genome_gifts)

# Print top pairs
day_clusters_negative_dist %>% 
  filter(count>3) %>% 
  tt()
```

#### Clusters per treatment

```{r community_clusters_treatment_negative, comment="", message=FALSE, warning=FALSE}
treatment_clusters_negative <- c(TG1$communities_neg,
                                 TG2$communities_neg,
                                 TG3$communities_neg,
                                 TG4$communities_neg,
                                 TG5$communities_neg) %>%
  purrr::map(~ combn(.x, 2, simplify = FALSE)) %>%
  unlist(recursive = FALSE) %>% 
  purrr::map(~ sort(.x))

treatment_clusters_negative_dist <- compute_cluster_distances(treatment_clusters_negative, genome_gifts)

# Print top pairs  --  all pairs actually appear only once
treatment_clusters_negative_dist %>% 
  filter(count>1) %>% 
  tt()
```

#### Cluster combined

```{r community_clusters_all_negative, comment="", message=FALSE, warning=FALSE}
all_clusters_negative <- c(day7$communities_neg,
                           day14$communities_neg,
                           day21$communities_neg,
                           day28$communities_neg,
                           day35$communities_neg,
                           TG1$communities_neg,
                           TG2$communities_neg,
                           TG3$communities_neg,
                           TG4$communities_neg,
                           TG5$communities_neg) %>%
  purrr::map(~ combn(.x, 2, simplify = FALSE)) %>%
  unlist(recursive = FALSE) %>% 
  purrr::map(~ sort(.x))

# Calculate functional distances
all_clusters_negative_dist <- compute_cluster_distances(all_clusters_negative, genome_gifts)

# Print top pairs
all_clusters_negative_dist %>% 
  filter(count>1) %>% 
  tt()
```

```{r community_clusters_all_plot2, comment="", message=FALSE, warning=FALSE}
day_clusters_negative_filtered <- day_clusters_negative_dist %>% 
  filter(count>3) %>%
  rename(count_day=count) %>% 
  mutate(clustering_day=TRUE)

treatment_clusters_negative_filtered <- treatment_clusters_negative_dist %>% 
  filter(count>3) %>%
  rename(count_treatment=count) %>% 
  mutate(clustering_treatment=TRUE)

all_clusters_negative_filtered <- all_clusters_negative_dist %>% 
  filter(count > 5) %>%
  rename(count_all=count) %>% 
  mutate(clustering_all=TRUE)

associated_pairs_negative <- all_clusters_negative_filtered %>% 
  full_join(day_clusters_negative_filtered, by=join_by("genomes"=="genomes","distance"=="distance")) %>% 
  full_join(treatment_clusters_negative_filtered, by=join_by("genomes"=="genomes","distance"=="distance")) %>%
  rowwise() %>% 
  mutate(count=max(count_all,count_day,count_treatment, na.rm=TRUE)) %>%
  mutate(clustering = case_when(
    clustering_all == TRUE ~ "all",  # If clustering_all is TRUE
    clustering_day == TRUE & is.na(clustering_all) & is.na(clustering_treatment) ~ "day", 
    clustering_treatment == TRUE & is.na(clustering_all) & is.na(clustering_day) ~ "treatment")) %>%
  select(genomes,count,clustering,distance) %>%
  mutate(clustering=factor(clustering)) %>%
  #add coordinates
  separate(genomes, into = c("genome1", "genome2"), sep = " - ") %>%
  left_join(gift_pcoa_vectors %>% rownames_to_column(var = "genome"), by = c("genome1" = "genome")) %>%
  rename(x1 = Axis.1, y1 = Axis.2) %>%
  left_join(gift_pcoa_vectors %>% rownames_to_column(var = "genome"), by = c("genome2" = "genome")) %>%
  rename(x2 = Axis.1, y2 = Axis.2)

associated_pairs_negative %>% 
  select(genome1,genome2,count,clustering) %>% 
  left_join(genome_metadata %>% select(genome,order,species),by=join_by(genome1==genome)) %>% 
  left_join(genome_metadata %>% select(genome,order,species),by=join_by(genome2==genome)) %>% 
  tt()
```

### Association vs. functional distance

```{r association_distance_pos_neg_test, comment="", message=FALSE, warning=FALSE}
weighted_sd <- function(x, w) {
  weighted_mean <- sum(w * x) / sum(w)
  variance <- sum(w * (x - weighted_mean)^2) / sum(w)
  sqrt(variance)
}

# Summary statistics
distance_stats <- bind_rows(associated_pairs_positive %>% mutate(type="positive"),
          associated_pairs_negative %>% mutate(type="negative")) %>% 
          left_join(genome_metadata %>% select(genome,genus),by=join_by(genome1==genome)) %>% 
          left_join(genome_metadata %>% select(genome,genus),by=join_by(genome2==genome)) %>%
          filter(genus.x!=genus.y) %>% #remove relationships with same genus due to crossmaping error probability
          group_by(type) %>% 
          summarise(weighted_mean = sum(distance * count) / sum(count),
                    weighed_sd = weighted_sd(distance, count))

distance_stats %>% 
  tt()

# Permutation test
weighted_mean <- function(x, w) {
  sum(x * w) / sum(w)
}

permutation_test <- function(data, n_permutations = 1000) {
  original_diff <- abs(weighted_mean(data$distance[data$type == "positive"], data$count[data$type == "positive"]) -
                         weighted_mean(data$distance[data$type == "negative"], data$count[data$type == "negative"]))
  
  perm_diff <- replicate(n_permutations, {
    shuffled_type <- sample(data$type)
    abs(weighted_mean(data$distance[shuffled_type == "positive"], data$count[shuffled_type == "positive"]) -
          weighted_mean(data$distance[shuffled_type == "negative"], data$count[shuffled_type == "negative"]))
  })
  
  p_value <- mean(perm_diff >= original_diff)
  return(p_value)
}

permutation_test(bind_rows(associated_pairs_positive %>% mutate(type="positive"),
                                      associated_pairs_negative %>% mutate(type="negative")), 
                n_permutations = 10000)

```

```{r association_distance_pos_background_test, comment="", message=FALSE, warning=FALSE}

# Calculate the weighted mean distance for the "positive" group
weighted_mean_positive <- sum(associated_pairs_positive$distance * associated_pairs_positive$count) / sum(associated_pairs_positive$count)

# Calculate the mean of all distances from the functional distance matrix
all_distances <- functional_distances[upper.tri(functional_distances, diag = FALSE)]
all_distances <- all_distances[!is.na(all_distances)]
sampled_distances <- sample(all_distances, size = 37, replace = FALSE)
mean_all_distances <- mean(sampled_distances)

# Number of permutations
n_permutations <- 10000

# Combine the positive distances and all distances into a single vector
combined_distances <- c(associated_pairs_positive$distance, sampled_distances)

# Get the observed difference in means
observed_diff <- weighted_mean_positive - mean_all_distances

# Permutation test
perm_diff <- replicate(n_permutations, {
  # Randomly shuffle the combined distances
  shuffled_distances <- sample(combined_distances)
  
  # Split into "positive" group (same size as original positive group) and "all" group
  shuffled_positive <- shuffled_distances[1:length(associated_pairs_positive$distance)]
  shuffled_all <- shuffled_distances[(length(associated_pairs_positive$distance) + 1):length(combined_distances)]
  
  # Calculate weighted mean for the shuffled positive group (use original counts)
  shuffled_weighted_mean_positive <- sum(shuffled_positive * associated_pairs_positive$count) / sum(associated_pairs_positive$count)
  
  # Calculate the mean for the shuffled all group
  shuffled_mean_all <- mean(shuffled_all)
  
  # Return the difference in means
  shuffled_weighted_mean_positive - shuffled_mean_all
})

# Calculate p-value (two-tailed test)
p_value <- mean(abs(perm_diff) >= abs(observed_diff))
p_value


```

```{r association_distance_neg_background_test, comment="", message=FALSE, warning=FALSE}
# Calculate the weighted mean distance for the "negative" group
weighted_mean_negative <- sum(associated_pairs_negative$distance * associated_pairs_negative$count) / sum(associated_pairs_negative$count)

# Calculate the mean of all distances from the functional distance matrix
all_distances <- functional_distances[upper.tri(functional_distances, diag = FALSE)]
all_distances <- all_distances[!is.na(all_distances)]
sampled_distances <- sample(all_distances, size = 37, replace = FALSE)
mean_all_distances <- mean(sampled_distances)

# Number of permutations
n_permutations <- 10000

# Combine the negative distances and all distances into a single vector
combined_distances <- c(associated_pairs_negative$distance, sampled_distances)

# Get the observed difference in means
observed_diff <- weighted_mean_negative - mean_all_distances

# Permutation test
perm_diff <- replicate(n_permutations, {
  # Randomly shuffle the combined distances
  shuffled_distances <- sample(combined_distances)
  
  # Split into "negative" group (same size as original negative group) and "all" group
  shuffled_negative <- shuffled_distances[1:length(associated_pairs_negative$distance)]
  shuffled_all <- shuffled_distances[(length(associated_pairs_negative$distance) + 1):length(combined_distances)]
  
  # Calculate weighted mean for the shuffled negative group (use original counts)
  shuffled_weighted_mean_negative <- sum(shuffled_negative * associated_pairs_negative$count) / sum(associated_pairs_negative$count)
  
  # Calculate the mean for the shuffled all group
  shuffled_mean_all <- mean(shuffled_all)
  
  # Return the difference in means
  shuffled_weighted_mean_negative - shuffled_mean_all
})

# Calculate p-value (two-tailed test)
p_value <- mean(abs(perm_diff) >= abs(observed_diff))
p_value


```

```{r association_distance_plot, comment="", message=FALSE, warning=FALSE}
tibble(dist=functional_distances[upper.tri(functional_distances, diag = FALSE)]) %>% 
  filter(!is.na(dist)) %>% 
  ggplot(aes(x = dist)) +
  geom_histogram(bins = 10, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_vline(xintercept = pull(distance_stats[1,2]), linetype="dotted", color = "red", size=1.5) +
  geom_vline(xintercept = pull(distance_stats[2,2]), linetype="dotted", color = "green", size=1.5) +
  theme_minimal()
```

### Persistent associations

```{r community_clusters_all_combined_plot, comment="", message=FALSE, warning=FALSE, fig.height=6, fig.width=10, fig.fullwidth=TRUE}
gift_pcoa_vectors %>%
  rownames_to_column(var="genome") %>%
  left_join(genome_metadata, by="genome") %>%
  ggplot() +
    geom_segment(data = associated_pairs_negative, 
               aes(x = x1, y = y1, xend = x2, yend = y2, linetype=clustering),
               color="#f5a2af",
               linewidth=1) +
    geom_segment(data = associated_pairs_positive, 
               aes(x = x1, y = y1, xend = x2, yend = y2, linetype=clustering),
               color="#bdf5a2",
               linewidth=1) +
    scale_linetype_manual(values=c("solid", "dashed", "dotted")) +
    geom_point(aes(x = Axis.1, y = Axis.2, color = order, size = length), 
             alpha = 0.9, shape = 16) +
    scale_color_manual(values = order_colors) +
    theme_minimal() + 
    theme(legend.position = "none")

```

```{r relationship_complementarities1, comment="", message=FALSE, warning=FALSE, fig.height=14, fig.width=10, fig.fullwidth=TRUE}
selected_genomes <- c("GPB:bin_000009","GPB:bin_000021","GPB:bin_000208","D300452:bin_000016","D300444:bin_000010","GPB:bin_000140")

#Functional difference matrix
genome_gifts[selected_genomes,] %>%
    to.elements(., GIFT_db) %>%
    as.data.frame() %>%
    stats::dist(., method = "manhattan") / ncol(genome_gifts[selected_genomes, ])

genome_gifts[selected_genomes, ] %>% 
  to.elements(.,GIFT_db=GIFT_db) %>%
  as.data.frame() %>% 
  rownames_to_column(var="genome") %>% 
  pivot_longer(!genome, names_to = "trait", values_to = "gift") %>%
  mutate(functionid = substr(trait, 1, 3)) %>%
    mutate(trait = case_when(
      trait %in% GIFT_db$Code_element ~ GIFT_db$Element[match(trait, GIFT_db$Code_element)],
      TRUE ~ trait
    )) %>%
    filter(!functionid %in% c("D04","D08","B09","B10","S01","S02","S03")) %>% 
    mutate(functionid = case_when(
      functionid %in% GIFT_db$Code_function ~ GIFT_db$Function[match(functionid, GIFT_db$Code_function)],
      TRUE ~ functionid
    )) %>%
    mutate(genome=factor(genome,levels=selected_genomes)) %>% 
    mutate(trait=factor(trait,levels=unique(GIFT_db$Element))) %>%
    mutate(functionid=factor(functionid,levels=unique(GIFT_db$Function))) %>%
    ggplot(aes(x=genome,y=trait,fill=gift)) +
        geom_tile(colour="white", linewidth=0.2)+
        scale_fill_gradientn(colours=rev(c("#d53e4f", "#f46d43", "#fdae61", "#fee08b", "#e6f598", "#abdda4", "#ddf1da")))+
        facet_grid(functionid ~ ., scales="free",space="free") +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
              strip.text.y = element_text(angle = 0)) + 
        labs(y="Traits",x="Samples",fill="GIFT")
```

```{r relationship_complementarities2, comment="", message=FALSE, warning=FALSE, fig.height=14, fig.width=10, fig.fullwidth=TRUE}
selected_genomes <- c("GPB:bin_000208","D300452:bin_000016","TG5_28:bin_000003","D300509:bin_000001","GPB:bin_000123","GPB:bin_000124","GPB:bin_000194")

#Functional difference matrix
genome_gifts[selected_genomes,] %>%
    to.elements(., GIFT_db) %>%
    as.data.frame() %>%
    stats::dist(., method = "manhattan") / ncol(genome_gifts[selected_genomes, ])

genome_gifts[selected_genomes, ] %>% 
  to.elements(.,GIFT_db=GIFT_db) %>%
  as.data.frame() %>% 
  rownames_to_column(var="genome") %>% 
  pivot_longer(!genome, names_to = "trait", values_to = "gift") %>%
  mutate(functionid = substr(trait, 1, 3)) %>%
    mutate(trait = case_when(
      trait %in% GIFT_db$Code_element ~ GIFT_db$Element[match(trait, GIFT_db$Code_element)],
      TRUE ~ trait
    )) %>%
    filter(!functionid %in% c("D04","D08","B09","B10","S01","S02","S03")) %>% 
    mutate(functionid = case_when(
      functionid %in% GIFT_db$Code_function ~ GIFT_db$Function[match(functionid, GIFT_db$Code_function)],
      TRUE ~ functionid
    )) %>%
    mutate(genome=factor(genome,levels=selected_genomes)) %>% 
    mutate(trait=factor(trait,levels=unique(GIFT_db$Element))) %>%
    mutate(functionid=factor(functionid,levels=unique(GIFT_db$Function))) %>%
    ggplot(aes(x=genome,y=trait,fill=gift)) +
        geom_tile(colour="white", linewidth=0.2)+
        scale_fill_gradientn(colours=rev(c("#d53e4f", "#f46d43", "#fdae61", "#fee08b", "#e6f598", "#abdda4", "#ddf1da")))+
        facet_grid(functionid ~ ., scales="free",space="free") +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
              strip.text.y = element_text(angle = 0)) + 
        labs(y="Traits",x="Samples",fill="GIFT")
```

