---
title: "3D'omics | Poultry Salmonella trial"
subtitle: "Manuscript in prep"
author:
  - Antton Alberdi^[University of Copenhagen, antton.alberdi@sund.ku.dk]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
url: https://3d-omics.github.io/AMAC005_chicken_salmonella
description: |
  Data analysis code for the study on the effect of probiotics on Salmonella infection in broiler chicken.
link-citations: yes
github-repo: 3d-omics/AMAC005_chicken_salmonella
---

```{r knitr_opts, echo=FALSE}
knitr::opts_chunk$set(
    class.source = "script-source",
    class.output = "script-output",
    comment = NA)
```

# Introduction

This webbook contains all the code used for the study on the effect of probiotics on Salmonella infection in broiler chicken.

## Prepare the R environment

### Environment

To reproduce all the analyses locally, clone this repository in your computer using:

```
RStudio > New Project > Version Control > Git
```

And indicating the following git repository:

> https://github.com/3d-omics/AMAC005_chicken_salmonella.git

Once the R project has been created, follow the instructions and code chunks shown in this webbook.

### Libraries

The following R packages are required for the data analysis.

```{r load_libraries, warning=FALSE, comments="", message=FALSE}
# Base
library(R.utils)
library(knitr)
library(tidyverse)
library(devtools)
library(tinytable)
library(broom)
library(broom.mixed)
library(rlang)


# For tree handling
library(ape)
library(phyloseq)
library(phytools)

# For plotting
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(ggnewscale)
library(ggtreeExtra)
library(ggtree)
library(ggh4x)
library(grid)
library(gridExtra)
library(corrplot)
library(png)

# For statistics
library(microbiome)
library(spaa)
library(vegan)
library(Rtsne)
library(geiger)
library(hilldiv2)
library(distillR)
library(ANCOMBC)
library(lme4)
library(Hmsc)
library(broom) 

# For networks
library(RCX)
library(RCy3)
library(igraph)
library(tibble)
library(scales)
library(dplyr)
```
